Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-11-09 RMS/final-version/TXT_15-TIE-3480.tex
finished TXT.tex

%\documentclass[onecolumn]{IEEEtranTIE}
\documentclass[journal]{IEEEtranTIE}
\usepackage{graphicx}
\usepackage[noadjust]{cite}
\usepackage{picinpar}
\usepackage{amsmath}
\usepackage{stfloats}
\usepackage{url}
\usepackage{flushend}
\usepackage[latin1]{inputenc}
\usepackage{colortbl}
\usepackage{soul}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{color}
\usepackage{alltt}
\usepackage[hidelinks,bookmarks=false]{hyperref}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{breakurl}
\usepackage{epstopdf}
\usepackage{pbox}

@@

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} \mathit{statePattern} \texttt{ |= } \mathit{prop} \texttt{ = } b
\texttt{ if } cond$, with $b$ a term of sort \verb|Bool|, stating that
$\mathit{prop}$ evaluates to $b$ in states which are instances of
$\mathit{statePattern}$ provided the condition $cond$ holds. These
equations together define $\mathit{prop}$ to hold in all states $s$
that make $s \texttt{ |= } \mathit{prop}$ evaluate to \verb|true|. A
temporal logic \emph{formula} is constructed by state propositions and
temporal logic operators such as \verb|~|(negation),
\verb|\/|(disjunction), \verb|[]|(``always''),
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

@@

Several auxiliary lemmas are proved:
\begin{lemma}
\label{l:auxtask}
Given $\mathit{ID}$ of sort \verb|MaybeNat| and $L$ of sort
\verb|TaskList|, $\mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
= \mteTask{\mathit{ID}}{L}\; \verb|monus|\;r$ for all
$r\le\mteTask{\mathit{ID}}{L}$.
\end{lemma}
\begin{IEEEproof}
If $\mathit{ID}=\verb|none|$, the case is trivial. By definition,
$\mteTask{\texttt{none}}{\deltaTask{\texttt{none}}{L}{r}} =
\mteTask{\texttt{none}}{L} = \verb|INF| =
\mteTask{\texttt{none}}{L}\\ \verb|monus|\; r$. Otherwise,
$\mathit{ID}= \verb|some|~N$ with $N$ of sort \verb|Nat|. Assume that
the \verb|cnt| of the $N$th task in $L$ is
\verb|[|$r_e$\verb|/|$C$\verb|]|. Then by definition,
$\deltaTask{\mathit{ID}}{L}{r}=L'$, where the
$N$th task in $L'$ has \verb|cnt| value being
\verb|[|$(r_e+r)$\verb|/|$C$\verb|]|. Hence,
\begin{IEEEeqnarray*}{Cl}
  & \mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
\\  
= & \mteTask{\mathit{ID}}{L'}
\\
= & C~\verb|monus|~(r_e+r) 
= (C~\verb|monus|~r_e)~\verb|monus|~r 
\\
= & \mteTask{\mathit{ID}}{L}~\verb|monus|~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxis}
Given $\mathit{ISRC}$ of class \verb|IntSrc| representing an
reasonable state of interrupt source in our model,
$\mteIS{\deltaIS{\mathit{ISRC}}{r}} = \mteIS{\mathit{ISRC}}\;
\verb|monus|\\ r$ for all $r\le\mteIS{\mathit{ISRC}}$.
\end{lemma}
\begin{IEEEproof}
A reasonable $\mathit{ISRC}$ must be of the form
$\IntSrc{O}{v}{T}$ with $v\le T$. Thus,
\begin{IEEEeqnarray*}{Cl}
  & \mteIS{\deltaIS{\mathit{ISRC}}{r}}
\\  
= & \mteIS{\IntSrc{O}{(v~\texttt{monus}~r)}{T}}
\\
= & v~\texttt{monus}~r
= \mteIS{\mathit{ISRC}}~\texttt{monus}~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxhw}
Given $\mathit{HW}$ of sort \verb|Hardware|, for all
$r\le\mteIr{\mathit{HW}}$, $\mteIr{\mathit{HW}} =
\mteIr{\mathit{HW}}\; \verb|monus|\; r$.
\end{lemma}
\begin{IEEEproof}
  It concludes by discussing on whether there exists an interrupt
  request detected in $\mathit{HW}$.
\end{IEEEproof}

@@

Condition~(iii). We must prove $\deltaS{\deltaS{s}{r}}{r'}\\ =
\deltaS{s}{r+r'}$ for all $r+r'\le\mteS{s}$, with $s$ being a system
state of sort \verb|System|. Using the same notations as in~(i), we
only consider the case $\mathit{ID}$\verb|::MaybeNat| in detail. By
definition, the left side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{\deltaS{s}{r}}{r'}
\\  
= & \verb|(|\deltaTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}{r'}
\\
  & \verb| |~T~\mathit{STS}~\mathit{HW}~\deltaIS{\deltaIS{\mathit{ISRC}}{r}}{r'}\verb|)|~\mbox{,}
\end{IEEEeqnarray*}
and the right side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{s}{r+r'}
\\  
= & \verb|(|\deltaTask{\mathit{ID}}{L}{r+r'}
\\
  & \verb| |~T~\mathit{STS}~\mathit{HW}~\deltaIS{\mathit{ISRC}}{r+r'}\verb|)|~\mbox{.}
\end{IEEEeqnarray*}
$\deltaTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}{r'}
=\deltaTask{\mathit{ID}}{L}{r+r'}$ holds thanks to the associativity
of $+$, while
$\deltaIS{\deltaIS{\mathit{ISRC}}{r}}{r'}\!=\!\deltaIS{\mathit{ISRC}}{r+r'}$
holds since $(v\; \verb|monus|\; r)\; \verb|monus|\; r'\\ = v\;
\verb|monus|\; (r+r')$ with $v$ of sort \verb|Time|. Thus (iii) holds.

@@

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_MIN_ZHOU.jpeg}}]%
{Min Zhou} received his B.S.\ degree in mathematics, and the
Ph.D.\ degree in computer science from Tsinghua University, Beijing,
China in 2007 and 2014, respectively.

@@

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_XIAOYU_SONG.jpeg}}]%
{Xiaoyu Song} received the Ph.D.\ degree from the University of Pisa,
Italy, in 1991. From 1992 to 1998, he was on the faculty at the
University of Montreal, Canada. He joined the Department of Electrical
and Computer Engineering at Portland State University in 1998, where
he is now a professor. He was an editor of IEEE Transactions on VLSI
Systems and IEEE Transactions on Circuits and Systems. He was awarded
an Intel Faculty Fellowship from 2000 to 2005. His research interests
include formal methods, design automation, embedded systems and
emerging technologies.
\end{IEEEbiography}

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_MING_GU.jpeg}}]%
{Ming Gu} received the B.S.\ degree in computer science from the
National University of Defense Technology, China in 1984, and the
M.S.\ degree in computer science from the Chinese Academy of Science,
China in 1986.

@@

\vspace{-5mm}
\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_JIAGUANG_SUN.jpeg}}]%
{Jiaguang Sun} received the B.S.\ degree in automation science from
Tsinghua University, China in 1970.

>>>>

\documentclass[12pt,onecolumn]{IEEEtranTIE}
%%\documentclass[journal]{IEEEtranTIE}
\usepackage{graphicx}
\usepackage[noadjust]{cite}
\usepackage{picinpar}
\usepackage{amsmath}
\usepackage{stfloats}
\usepackage{url}
\usepackage{flushend}
\usepackage[latin1]{inputenc}
\usepackage{colortbl}
\usepackage{soul}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{color}
\usepackage{alltt}
\usepackage[hidelinks,bookmarks=false]{hyperref}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{breakurl}
\usepackage{epstopdf}
\usepackage{pbox}

@@

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form\\ $\texttt{ceq
} \mathit{statePattern} \texttt{ |= } \mathit{prop} \texttt{ = } b
\texttt{ if } cond$, with $b$ a term of sort \verb|Bool|, stating that
$\mathit{prop}$ evaluates to $b$ in states which are instances of
$\mathit{statePattern}$ provided the condition $cond$ holds. These
equations together define $\mathit{prop}$ to hold in all states $s$
that make $s \texttt{ |= } \mathit{prop}$ evaluate to \verb|true|. A
temporal logic \emph{formula} is constructed by state propositions and
temporal logic operators such as \verb|~|(negation),
\verb|\/|(disjunction), \verb|[]|(``always''),
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

@@

Several auxiliary lemmas are proved:
\begin{lemma}
\label{l:auxtask}
Given $\mathit{ID}$ of sort \verb|MaybeNat| and $L$ of sort
\verb|TaskList|,\\ $\mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
= \mteTask{\mathit{ID}}{L}\; \verb|monus|\;r$ for all
$r\le\mteTask{\mathit{ID}}{L}$.
\end{lemma}
\begin{IEEEproof}
If $\mathit{ID}=\verb|none|$, the case is trivial. By definition,
$\mteTask{\texttt{none}}{\deltaTask{\texttt{none}}{L}{r}}\\ =
\mteTask{\texttt{none}}{L} = \verb|INF| =
\mteTask{\texttt{none}}{L}\; \verb|monus|\; r$. Otherwise,
$\mathit{ID}= \verb|some|~N$ with $N$ of sort \verb|Nat|. Assume that
the \verb|cnt| of the $N$th task in $L$ is
\verb|[|$r_e$\verb|/|$C$\verb|]|. Then by definition,
$\deltaTask{\mathit{ID}}{L}{r}\\ = L'$, where the
$N$th task in $L'$ has \verb|cnt| value being
\verb|[|$(r_e+r)$\verb|/|$C$\verb|]|. Hence,
\begin{IEEEeqnarray*}{Cl}
  & \mteTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}
\\  
= & \mteTask{\mathit{ID}}{L'}
\\
= & C~\verb|monus|~(r_e+r) 
= (C~\verb|monus|~r_e)~\verb|monus|~r 
\\
= & \mteTask{\mathit{ID}}{L}~\verb|monus|~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxis}
Given $\mathit{ISRC}$ of class \verb|IntSrc| representing an
reasonable state of interrupt source in our model,
$\mteIS{\deltaIS{\mathit{ISRC}}{r}} = \mteIS{\mathit{ISRC}}\;
\verb|monus|\; r$ for all $r\le\mteIS{\mathit{ISRC}}$.
\end{lemma}
\begin{IEEEproof}
A reasonable $\mathit{ISRC}$ must be of the form
$\IntSrc{O}{v}{T}$ with $v\le T$. Thus,
\begin{IEEEeqnarray*}{Cl}
  & \mteIS{\deltaIS{\mathit{ISRC}}{r}}
\\  
= & \mteIS{\IntSrc{O}{(v~\texttt{monus}~r)}{T}}
\\
= & v~\texttt{monus}~r
= \mteIS{\mathit{ISRC}}~\texttt{monus}~r\;. \hfill\IEEEQEDhere
\end{IEEEeqnarray*}
\end{IEEEproof}

\begin{lemma}
\label{l:auxhw}
Given $\mathit{HW}$ of sort \verb|Hardware|, for all
$r\le\mteIr{\mathit{HW}}$,\\ $\mteIr{\mathit{HW}} =
\mteIr{\mathit{HW}}\; \verb|monus|\; r$.
\end{lemma}
\begin{IEEEproof}
  It concludes by discussing on whether there exists an interrupt
  request detected in $\mathit{HW}$.
\end{IEEEproof}

@@

Condition~(iii). We must prove $\deltaS{\deltaS{s}{r}}{r'} =
\deltaS{s}{r+r'}$ for all $r+r'\le\mteS{s}$, with $s$ being a system
state of sort \verb|System|. Using the same notations as in~(i), we
only consider the case $\mathit{ID}$\verb|::MaybeNat| in detail. By
definition, the left side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{\deltaS{s}{r}}{r'}
\\  
= & \verb|(|\deltaTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}{r'}
\\
  & \verb| |~T~\mathit{STS}~\mathit{HW}~\deltaIS{\deltaIS{\mathit{ISRC}}{r}}{r'}\verb|)|~\mbox{,}
\end{IEEEeqnarray*}
and the right side of the equation
\begin{IEEEeqnarray*}{Cl}
  & \deltaS{s}{r+r'}
\\  
= & \verb|(|\deltaTask{\mathit{ID}}{L}{r+r'}
\\
  & \verb| |~T~\mathit{STS}~\mathit{HW}~\deltaIS{\mathit{ISRC}}{r+r'}\verb|)|~\mbox{.}
\end{IEEEeqnarray*}
$\deltaTask{\mathit{ID}}{\deltaTask{\mathit{ID}}{L}{r}}{r'}
=\deltaTask{\mathit{ID}}{L}{r+r'}$ holds thanks to the associativity
of $+$, while
$\deltaIS{\deltaIS{\mathit{ISRC}}{r}}{r'}=\deltaIS{\mathit{ISRC}}{r+r'}$
holds since $(v\; \verb|monus|\; r)\; \verb|monus|\; r' = v\;
\verb|monus|\; (r+r')$ with $v$ of sort \verb|Time|. Thus (iii) holds.

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_MIN_ZHOU.jpeg}}]%
{Min Zhou} received his B.S.\ degree in mathematics, and the
Ph.D.\ degree in computer science from Tsinghua University, Beijing,
China in 2007 and 2014, respectively.

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_XIAOYU_SONG.jpeg}}]%
{Xiaoyu Song} received the Ph.D.\ degree from the University of Pisa,
Italy, in 1991. From 1992 to 1998, he was on the faculty at the
University of Montreal, Canada. He joined the Department of Electrical
and Computer Engineering at Portland State University in 1998, where
he is now a professor. He was an editor of IEEE Transactions on VLSI
Systems and IEEE Transactions on Circuits and Systems. He was awarded
an Intel Faculty Fellowship from 2000 to 2005. His research interests
include formal methods, design automation, embedded systems and
emerging technologies.
\end{IEEEbiography}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_MING_GU.jpeg}}]%
{Ming Gu} received the B.S.\ degree in computer science from the
National University of Defense Technology, China in 1984, and the
M.S.\ degree in computer science from the Chinese Academy of Science,
China in 1986.

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{AUTHOR_JIAGUANG_SUN.jpeg}}]%
{Jiaguang Sun} received the B.S.\ degree in automation science from
Tsinghua University, China in 1970.
